### 2020-04-02 llncs2e/figure2.tex

\begin{figure}[t!]
\centering
\begin{tikzcd}
& M\simeq \mathbb{R}^n \arrow[ld, "x^{(0)}", bend right = 15] 
& U \simeq \mathbb{R}^{n-k} \arrow[ld, "x^{(1)}" pos=0.8, swap, bend right = 15] \arrow[d, "x^{(2)}"] \arrow[rd, "x^{(L-1)}", bend left = 15] \arrow[l, hook', swap, "\imath"] 
&\\
x^{(0)}(M) \arrow[r, "\varphi^{(0)}"] 
& x^{(1)}(U) \arrow[r, leftrightarrow, "\varphi^{(1)}"]
& x^{(2)}(U) \arrow[r, leftrightarrow, "\varphi^{(L-1)}", dotted]
& x^{(L-1)}(U) \arrow[r, hook, "\varphi^{(L)}"]
& x^{(L)}(M)
\end{tikzcd}
\caption{The architecture of the neural net maps over $\varphi^{(0)}$ into a subspace of $M$, in
which the searched manifold $U$ of the data can be embedded. The maps $\varphi^{(1)}, \cdots, \varphi^{(L-1)}$
are models of bijective, smooth and therefore diffeomorphic functions.}
\label{architecture}
\end{figure}