### 2020-04-02 llncs2e/figure3.tex

\begin{figure}[t!]
\begin{tikzpicture}
\begin{axis}[
	height=2.3in,
	width= 2.7in,
	axis x line = bottom,
	axis y line = left,
	minor x tick num=10,
	ylabel={$\frac{\omega_i + \omega_{j+1}}{2}$},
	xlabel={(a) cifar10.},
	xmin=0,
	xmax=1050,
	ymin=0,
	ymax=650,
	ticklabel style={font=\small},
	x label style={at={(axis description cs:0.5,-0.1)},anchor=north},
	y label style={at={(axis description cs:0.20,0.85)},rotate=-90,anchor=south, fill=white},
	axis line style={-{Latex[length=1.5mm,width=1.5mm]}}
]
\coordinate (spypoint) at (axis cs:85,50);
\coordinate (magnifyglass) at (axis cs:800,400);
\draw [thick, dotted, draw=gray] (axis cs: 0,580) -- (axis cs: 1050,580) node[pos=0.9, below] {$\infty$};
\end{axis}
\begin{scope}[spy using outlines={circle, magnification=4, connect spies}]
\node[inner sep=0pt] at (2.55,1.9) {\includegraphics[width=0.42\textwidth]{figure3_a.pdf}};
\spy [lightgray, size=2cm] on (spypoint) in node[fill=white] at (magnifyglass);
\end{scope} 
\end{tikzpicture}\hfill
\begin{tikzpicture}
\begin{axis}[
	height=2.3in,
	width= 2.7in,
	axis x line = bottom,
	axis y line = left,
	minor x tick num=10,
	xlabel={(b) cifar100.},
	xmin=0,
	xmax=1050,
	ylabel={$\cdot 10^{3}$},
	ymin=0,
	ymax=5,
	ticklabel style={font=\small, fill=white},
	x label style={at={(axis description cs:0.5,-0.1)},anchor=north},
	y label style={at={(axis description cs:0.1,0.9)},rotate=-90,anchor=south},
	axis line style={-{Latex[length=1.5mm,width=1.5mm]}}
]
\draw [thick, dotted, draw=gray] (axis cs: 0,4.45) -- (axis cs: 1050,4.45) node[pos=0.9, below] {$\infty$};
\end{axis}
\node[inner sep=0pt] (h) at (2.55,1.9) {\includegraphics[width=0.42\textwidth]{figure3_b.pdf}};
\node[inner sep=0pt] (h) at (4.5,3.1) {{\color{blacky}$\bullet$} {$H_0$}};
\node[inner sep=0pt] (h) at (4.5,2.7) {{\color{gray}$\bullet$} {$H_1$}};
\node[inner sep=0pt] (h) at (4.5,2.3) {{\color{purpur}$\bullet$} {$H_2$}};
\node[inner sep=0pt] (h) at (4.5,1.9) {{\color{brick}$\bullet$} {$H_3$}};
\node[inner sep=0pt] (h) at (4.5,1.5) {{\color{fire}$\bullet$} {$H_4$}};
\end{tikzpicture}
\caption{Mean confident persistence landscapes of the investigated datasets cifar10 and cifar100 with a resolution of $1000$. The mean persistence landscapes were computed on truncated persistence diagrams, with a confidence of $\alpha = 0.95$ for a representative of a persistent homology class tracked on the filtration.}
\label{homgroups}
\end{figure}