### 2020-04-02 llncs2e/figure5.tex

\begin{table*}[t!]
\centering
\caption{The representatives from the mean confidence persistent landscape are counted per homology group. We give the minimal embedding dimension for some compact object on a connected commutative Lie group,
having approximately as many representatives of the respective homology groups. Further, we give an embedding dimension for an object consisting of all such representatives. We apply Eq. \ref{statespace} to compute the required state space dimension of resnets with different degrees of smoothness.}
\begin{tabular}{@{}rccccccccccccc@{}}
\toprule \multirow{2}{*}{\backslashbox{Data \strut}{\strut Features}} & \phantom{abc} & \multicolumn{5}{c}{Homology groups} &&& \multicolumn{5}{c}{$\approx$ Embedding dim.}\\
\cmidrule{3-7} \cmidrule{10-14} & & $H_0$ & $H_1$ & $H_2$ & $H_3$ & $H_4$ &&& $U$ & $\mathcal{C}^1$ & $\mathcal{C}^2$ & $\mathcal{C}^3$ & $\mathcal{C}^4$ \\
\midrule
cifar10 && 12 & 16 & 40 & 59 & 50 &&& $184 \pm 44$ & x & x & x & x\\
cifar100 && 13 & 18 & 34 & 46 & 48 &&& $$ & x & x & x & x \\
\midrule
Solutions of && $12$ & $16$ & $9 \pm 4$ & $8 \pm 3$ & $7 \pm 15$\\
Eq. \ref{countingsimple} \& Eq. \ref{countingequation} && $13$ & $18$ & $9 \pm 2$ & $8 \pm 10$ & $7 \pm 13$\\
\cmidrule{1-8}
\end{tabular}
\label{stats}
\end{table*}