### 2020-06-17 llncs2e/figure1.tex
update paper

\begin{figure}[t]
    \begin{overpic}[width=\textwidth,tics=10]{figure1.pdf}
        \put (0.9,58) {\textbf{a})}
        \put (0.9,28) {\textbf{b})}
        \put (70,58) {\textbf{c})}
        \put (8.8,29.6) {$0$}
        \put (32.8,29.6) {$0$}
        \put (58,29.6) {$0$}
        \put (74.8,43.6) {\small{1}}
        \put (90.5,43.6) {\small{2}}
        \put (84.6,22.4) {\small{3}}
        \put (84.6,1) {\small{4}}
        \put (7,13) {$K_1$}
        \put (31,13) {$K_2$}
        \put (57,13) {$K_3$}
        \put (21,13) {$\subseteq$}
        \put (45,13) {$\subseteq$}
        \put (32,39) {$B_k$}
        \put (32,49) {$Z_k$}
        \put (32,55) {$C_k$}
        \put (56,39) {$B_{k-1}$}
        \put (56,49) {$Z_{k-1}$}
        \put (56,55) {$C_{k-1}$}
        \put (7,39) {$B_{k+1}$}
        \put (7,49) {$Z_{k+1}$}
        \put (7,55) {$C_{k+1}$}
        \put (19.5,56) {$\partial_{k+1}$}
        \put (44.5,56) {$\partial_{k}$}
        \put (92,37) {{\color{red5} $\bullet$} $H_0$}
        \put (92,34) {{\color{red4} $\bullet$} $H_1$}
        \put (92,31) {{\color{red3} $\bullet$} $H_2$}
        \put (92,28) {{\color{red2} $\bullet$} $H_3$}
        \put (92,25) {$\cdots$}
        \put (92,15.5) {{\color{red5} $\bullet$} $H_0$}
        \put (92,12.5) {{\color{red4} $\bullet$} $H_1$}
        \put (92,9.5) {{\color{red3} $\bullet$} $H_2$}
        \put (92,6.5) {{\color{red2} $\bullet$} $H_3$}
        \put (92,3.5) {$\cdots$}
    \end{overpic}\\[0.1cm]
    \caption{\textbf{a}) Illustration of the chain complex following Zomorodian et al. \cite{ZomorodianC05}. \textbf{b}) Sublevel sets of a simplicial complex connected by inclusion. \textbf{c}) Pipeline for persistent homology: \textbf{c})1. Loading point sets. \textbf{c})2. Computation of a filtration. \textbf{c})3. Computation of persistent homology with confidence band. \textbf{c})4. Computation of persistence landscapes.}
    \label{figure1}
\end{figure}