### 2020-06-17 llncs2e/figure3.tex

\begin{table*}[t!]
\centering
\caption{Counts of the representatives per homology group from $\text{PL}(X)$.}
\label{figure3}
\begin{tabular}{@{}rcccccccccccccc@{}}
\toprule \multirow{2}{*}{\backslashbox{\texttt{Data} \strut}{\strut \texttt{Features}}} & \phantom{abc} & \multicolumn{5}{c}{\texttt{Homology groups}} &&& \multicolumn{6}{c}{\texttt{$\approx$ embedding dimension}}\\
\cmidrule{3-7} \cmidrule{10-15} & & $H_0$ & $H_1$ & $H_2$ & $H_3$ & $H_4$ &&& $p$ & $q \vert {H_1}$ & $q \vert {H_2}$ & $q \vert {H_3}$ & $q \vert {H_4}$ & $\dim U$\\
\midrule
\texttt{cifar10} && 12 & 16 & 40 & 59 & 50 &&& $12$ & $16$ & $9 \scriptstyle{\pm 4}$ & $8 \scriptstyle{\pm 3}$ & $7 \scriptstyle{\pm 15}$ & $92 \scriptstyle{\pm 44}$\\
\texttt{cifar100} && 13 & 18 & 34 & 46 & 48 &&& $13$ & $18$ & $9 \scriptstyle{\pm 2}$ & $8 \scriptstyle{\pm 10}$ & $7 \scriptstyle{\pm 13}$ & $97 \scriptstyle{\pm 50}$\\
\midrule
\end{tabular}
\label{stats}
\end{table*}